### 2017-12-21 submission/manuscript.Rmd
Skeleton of manuscript.Rmd
Revert all commits to the content to do branch based writing with Pull Requests from @Sjakie37 and @Wicherts

---
title: 'Understanding data fabrication: Qualitative Comparative Analysis (QCA) of
fabrication strategies'
author: "CHJ Hartgerink"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  word_document: default
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_depth: 2
csl: ../bibliography/apa.csl
bibliography: ../bibliography/library.bib
---

# Abstract

# Introduction

# Methods

# Results

# Discussion

# Author's note

All materials used in this project are available at [https://github.com/chartgerink/2015ori-2](https://github.com/chartgerink/2015ori-2) and are preserved at Zenodo.

# References